Assessing the potential of Jacobian-free Newton-Krylov methods for cell-centred finite volume solid mechanics

### 2024-12-15 article.tex
Updates to the text: add appendix for material laws

%The two forms are connected through Nanson’s formula \cite{bathe_finite_1996}, which relate the deformed area vector $\bb{\Gamma}$ with the initial area vector $\bb{\Gamma}_{o}$:
%\begin{equation}
%    \bb{\Gamma} = J\bb{F}^{-T}\cdot\bb{\Gamma}_o
%\end{equation}

%Although the total Lagrangian approach is a viable option for wire drawing, the current work adopts the updated Lagrangian approach as developing Eulerian-type upstream and downstream conditions (Section \ref{sec:euler_BCs}) is conceptually easier in an updated Lagrangian formulation.

The definition of the engineering stress ($\bb{\sigma}_s$) and true stress ($\bb{\sigma}$) in Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL} is given by a chosen mechanical law, e.g. linear elasticity, neo-Hookean hyperelasticity.
Several mechanical laws are considered in this work, as briefly described in Section \ref{sec:test_cases}.

\hl{Mention BCs: disp, trac, symm}.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Cell-Centred Finite Volume Discretisation}
\label{sec:discretisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this work, a nominally second-order cell-centred finite volume discretisation is employed.
% based on previously described work, for example, \citep{Cardiff2017, Batistic2022, Tukovic2013, Jasak2000}.
%Consequently, only a summary of the discretisation is presented below.
The solution domain is discretised in both space and time.
The total simulation period is divided into a finite number of time increments, denoted as $\Delta t$, and the discretised governing momentum equation is solved iteratively in a time-marching fashion.
%The spatial domain is partitioned into a finite number of contiguous convex polyhedral cells.
The spatial domain is partitioned into a finite set $\mathcal{P}$ of contiguous convex polyhedral cells, where each cell is denoted by $P \in \mathcal{P}$.
A representative cell $P$ is shown in Figure~\ref{fig:cell}.
The set of all faces of cell $P$ is denoted by $\mathcal{F}_P$. This set is further subdivided into two disjoint subsets:
\begin{itemize}
    \item \textbf{Internal Faces} ($\mathcal{F}^{\text{int}}_P$): Faces that are shared with neighbouring cells.
    \item \textbf{Boundary Faces} ($\mathcal{F}^{\text{bnd}}_P$): Faces that lie on the boundary of the spatial domain.
    The boundary faces of cell $P$ are further classed into three disjoint sets, $\mathcal{F}^{\text{bnd}}_P \coloneqq \mathcal{F}^{\text{disp}}_P \cup \mathcal{F}^{\text{trac}}_P \cup \mathcal{F}^{\text{symm}}_P$, representing boundary faces where displacement ($\mathcal{F}^{\text{disp}}_P$), traction ($\mathcal{F}^{\text{trac}}_P$) and symmetry ($\mathcal{F}^{\text{symm}}_P$) conditions are prescribed.
       % \quad \text{with} \quad \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{trac}}_P = \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{symm}}_P = \mathcal{F}^{\text{trac}}_P \cap \mathcal{F}^{\text{symm}}_P = \emptyset.$
\end{itemize}
Each internal face $f_i \in \mathcal{F}^{\text{int}}_P$ corresponds to a neighbouring cell $N_{f_i} \in \mathcal{N}_P$, where $\mathcal{N}_P$ is the set of all neighbouring cells of $P$. The outward unit normal vector associated with an internal face $f_i$ is denoted by $\mathbf{n}_{f_i}$, while the outward unit normal vector associated with a boundary face $b_i$ is denoted by $\mathbf{n}_{b_i}$.
The vector $\mathbf{d}_{f_i}$ connects the centroid of cell $P$ with the centroid of the neighbouring cell $N_{f_i}$, whereas the vector $\mathbf{d}_{b_i}$ connects the centroid of cell $P$ with the centroid of boundary face ${b_i}$.
For convenience, we will also define the set of all faces of cell $P$ excluding those on a traction boundary as $\mathcal{F}_P^{\text{non-trac}} \coloneqq \mathcal{F}_P^{\text{int}} \cup \mathcal{F}_P^{\text{disp}} \cup \mathcal{F}_P^{\text{symm}}$.
%The proposed solution discretisation follows closely the approach of \citet{cardiff_lagrangian_2017}; consequently, only an overview of the final discretised form of equations and adopted solution algorithm are given below.
\begin{figure}[htbp]
	\centering
%	\subfigure[Magnitude of the manufactured displacement solution]
%	{
%		\label{fig:mms_solution}
   		\includegraphics[height=0.45\textwidth]{figures/cell} 
%   	}
%	\subfigure[Polyhedral mesh with $1\,000$ cells]
%	{
%		\label{fig:mms_mesh}
%   		\includegraphics[height=0.45\textwidth]{figures/mms_mesh}  
%   	}
	\caption{Representative convex polyhedral cell $P$ and neighbouring cell $N_{f_i}$, which share a face $f_i$. \hl{Add boundary face}}
	\label{fig:cell}
\end{figure}

@@

Using Equation \ref{eq:volume_integral}, the inertia term (e.g. left-hand side term of Equation \ref{eqn:momentum_lingeom}) becomes
\begin{eqnarray} \label{eq:inertia}
	\int_{\Omega_P} \rho \frac{\partial \bb{u} }{\partial t}  d \Omega_P
%	\;&\approx&\;
%	\int_{\mathrm{\Omega}} \rho \left[ \frac{\partial \bb{u} }{\partial t}_P + (\bb{x} - \bb{x}_P) \cdot \nabla \frac{\partial \bb{u} }{\partial t}_P \right] d\mathrm{\Omega} \notag \\
	\;&\approx&\;
	\rho_P \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P  \Omega_P
\end{eqnarray}
Similarly, the body force term (e.g. the second term on the right-hand side of Equation \ref{eqn:momentum_lingeom}) becomes:
\begin{eqnarray}
	\int_{\Omega_P} \, \rho \, \bb{g} \,  d \Omega_P
	\;&\approx&\;
	\rho_P \, \bb{g}\,  \Omega_P
\end{eqnarray}
%where subscript $P$ indicates a quantity at the cell centre.
The discretisation of the acceleration in Equation \ref{eq:inertia} can be achieved using one of many finite difference schemes, e.g. first-order Euler, second-order backwards, second-order Newmark-beta.
In the current work, the second-order backwards (BDF2) scheme is used:
%\begin{eqnarray}
%	 \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P
%	 &\approx& \frac{3\bb{v}_{t+1} - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t} \notag \\
%	&\approx& \frac{3\left( \frac{3\bb{u}_{t+1} - 4\bb{u}_{t} + \bb{u}_{t-1}}{2\Delta t} \right) - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t}
%\end{eqnarray}
\begin{eqnarray}
	\left(\frac{\partial^2 \boldsymbol{u}_P}{\partial t^2}\right)_P
	&\approx& \frac{3\boldsymbol{v}_P^{[t+1]} - 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t} \notag \\
	&\approx&
	\frac{3\left( 
		\frac{3\boldsymbol{u}_P^{[t+1]} - 4\boldsymbol{u}_P^{[t]} + \boldsymbol{u}_P^{[t-1]}}{2\Delta t} 
		\right) 
	- 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t}
\end{eqnarray}
where $\Delta t$ is the time increment -- assumed constant here -- and $\bb{v} = \partial \bb{u}/\partial t$ is the velocity vector.
Superscript $[t]$ indicates the time level, with $\bb{u}_P^{[t+1]}$ corresponding to the unknown displacement at the current time step.
Consequently, the displacement and velocity at the two previous time steps must be stored, or alternatively the displacement at the previous four time steps.

@@

@@

In Equation \ref{eq:RhieChow}, the first terms within the square brackets on the right-hand side represent a compact stencil (two-node) approximation of the face normal gradient, while the second terms represent a larger stencil approximation.
These two terms cancel out in the limit of mesh refinement (or if the solution varies linearly); otherwise, they produce a stabilisation effect that tends to smooth the solution fields.
As the term reduces at a third-order rate, it does not affect the overall scheme's second-order accuracy; the 1-D local truncation error is derived for this Rhie-Chow stabilisation term in Appendix \ref{app:RhieChow}.

\hl{HERE}

@@

\subsubsection{Boundary Conditions}
%Boundary faces $f_i \in \mathcal{F}_{\text{bnd},P}$ are classified based on boundary conditions into displacement-type, traction-type, and symmetry-type faces.
%\hl{Comment on traction boundaries} \hl{extrapolate to get value or use constitutive law}
%Boundary conditions must be applied to the faces that coincide with the boundary of the solution domain.
%The discretised expressions on boundary faces are modified to account for either the known displacement components in Dirichlet conditions or the known traction for Neumann conditions.
%In the current approach, boundary face 
Enforcing boundary conditions involves modifying the discretised surface integrals at the boundaries faces.
Three types of boundary condition are considered here: prescribed displacement, prescribed traction, and symmetry planes.

\hl{Include: When calculating cell-centre gradients in cells adjacent to the boundary, known boundary values are used for displacement boundaries, whereas traction boundary faces are excluded.}

Two terms must be approximated on a boundary face $b$:
\begin{eqnarray} \label{eq:BCtermsLap}
	\int_{\mathrm{\Gamma}_{b}} K_{imp} \bb{n}_{b} \cdot \left[ \nabla\left(\Delta\bb{u}\right) \right]_b d\mathrm{\Gamma}_{b}
\end{eqnarray}
which comes from Equation \ref{eq:diffusion}, and 
\begin{eqnarray} \label{eq:BCtermsStress}
	\int_{\mathrm{\Gamma_b}}(j_b \bb{f}^{-T}_b \cdot \bb{n}_b) \cdot \boldsymbol{\sigma}_b \ d\mathrm{\Gamma_b}
\end{eqnarray}
which comes from Equation \ref{eq:surfaceStress}, where subscript $b$ indicates a quantity at a boundary face in the updated configuration.

Equation \ref{eq:BCtermsLap} must be discretised both implicitly and explicitly (deferred correction), as per Equation \ref{eqn:MomentumImplicitExplicit}, while Equation \ref{eq:BCtermsStress} is discretised explicitly.
The Rhie-Chow term is set to zero on boundary faces, and all geometric terms ($\bb{n}_{b}$, $\mathrm{\Gamma}_{b}$) are calculated on the updated configuration.

Five classes of boundary conditions are employed in the current work (prescribed displacement, prescribed traction, symmetry plane, axisymmetric/wedge, frictional contact) as described below.
An additional boundary condition where the normal component of the displacement is prescribed and the shear traction is zero is also used; the discretisation of this condition employs the prescribed displacement condition in the boundary normal direction and the traction condition in the tangential direction.
%\begin{itemize}
%	\item Prescribed displacement,
%	\item Prescribed traction,
%	\item Symmetry,
%	\item Axisymmetric (wedge),
%	\item Frictional contact.
%\end{itemize}
%This enforcement of these five conditions is described below.

\paragraph{Prescribed displacement}
The displacement boundary condition, a Dirichlet condition, may be constant in time or time-varying and fixes the value of $\bb{u}$ at the centre of a boundary face.
The corresponding displacement increment at the boundary face $\Delta \bb{u}_b = \bb{u}^{[m+1]}_b - \bb{u}^{m}_b$ is substituted into the calculation of the surface normal gradient term in Equation \ref{eq:BCtermsLap}.
The resulting contribution of a boundary face $b$ to the discretised Laplacian term becomes:
\begin{eqnarray} \label{eq:diffusionBC}
%	\underbrace{
	\int_{\mathrm{\Gamma}_{b}}
	 K_{imp} \bb{n}_{b} \cdot \left[ \nabla\left(\Delta\bb{u}\right) \right]_b d\mathrm{\Gamma}_{b}
%	 }_{\text {implicit}}
	&\approx&
	K_{imp}^b \left|\boldsymbol{\Delta}_{b} \right|\left(\frac{\Delta \bb{u}_{b} - \Delta \bb{u}_P}{\left|\bb{d}_b\right|}\right)\left|\bb{\Gamma}_{b}\right| \notag \\
	    &&+ K_{imp}^b \; \bb{k}_{b} \cdot
	    \left[
	    \boldsymbol{\nabla} \left(\mathrm{\Delta}\bb{u}\right)
	    \right]_P
	    \left|\bb{\Gamma}_{b}\right|
\end{eqnarray}
where $\bb{d}_b$ connects the cell centre $\bb{C}_P$ with the boundary face centre $\bb{C}_b$ in the updated configuration, and the non-orthogonal correction vector $\bb{k}_{b} = (\textbf{I} - \bb{n}_b \bb{n}_b) \cdot \bb{d}_b$.

\paragraph{Prescribed traction}
The traction boundary condition, constant in time or time-varying, is implemented by replacing $\bb{n}_f \cdot \bb{\sigma}_f$ on boundary faces with the known prescribed traction $\bb{t}_b$.
Following calculation of the cell-centred displacements, displacements on traction boundary faces $\bb{u}_b$ is calculated by extrapolation from the adjacent cell centre:
\begin{eqnarray}
	\bb{u}_b = \bb{u}_P + (\bb{x}_b - \bb{x}_P) \cdot \left(\bb{\nabla} \bb{u} \right)_P
\end{eqnarray}
where subscript $b$ indicates a value at the centre of a boundary face, and $P$ indicates a value at the centre of the adjacent cell.

\paragraph{Symmetry plane}
On a symmetry plane normal component of displacement is zero, and the tangential gradient of displacement is zero.
For boundary faces on a symmetry plane, the mirror reflection $\left(\bb{\Delta} \bb{u}\right)_N$ of the cell-centre displacement increment $\left(\bb{\Delta} \bb{u}\right)_P$ can be obtained as
\begin{eqnarray}
	\left(\bb{\Delta} \bb{u}\right)_N &=& \bb{R}_m \cdot \left(\bb{\Delta} \bb{u}\right)_P \notag \\
		&=& \left(\textbf{I} - 2 \bb{n}_b \bb{n}_b \right) \cdot \left(\bb{\Delta} \bb{u}\right)_P
\end{eqnarray}
where $\bb{R}_m$ is termed the reflection tensor.
The discretisation in Equation \ref{eq:diffusion} takes the same form, but the neighbour cell-centre values are replaced by the mirror reflected values for boundary faces on the symmetry plane.
The $\bb{d}_b$ vector at a symmetry plane boundary face is calculated as
\begin{eqnarray}
	 \bb{d}_b &=& \bb{R}_m \cdot \bb{C}_P - \bb{C}_P
\end{eqnarray}
%where $\bb{C}_P$ is the positional vector of the centre of cell $P$.
This means that symmetry planes are orthogonal by definition, and the non-orthogonal correction (second term on the right-hand side of Equation \ref{eq:diffusion}) drops out.

In the current work, the cell-centre values are corrected for skewness errors, resulting in the following contribution to the Laplacian term at a symmetry boundary face (Equation \ref{eq:diffusionBC}):
\begin{eqnarray}
	\int_{\mathrm{\Gamma}_{u_b}}	 K_{imp} \bb{n}_{b} \cdot\nabla\left(\Delta\bb{u}\right) d\mathrm{\Gamma}_{b}
	&=&
	K_{imp}^b \left|\boldsymbol{\Delta}_{b} \right| \left( \frac{\bb{\Delta} \bb{u}_N^* - \Delta \bb{u}_P^*}{\left|\bb{d}_b\right|} \right) \left|\bb{\Gamma}_{b}\right|
\end{eqnarray}
where the skew-corrected cell-centre values are
\begin{eqnarray}
	 \bb{\Delta} \bb{u}_P^* &=&
	 	\bb{\Delta} \bb{u}_P + \left[ (\textbf{I} - \bb{n}_b \bb{n}_b) \cdot \bb{d}_b \right] \cdot \left[ \bb{\nabla} \left(\bb{\Delta} \bb{u}\right) \right]_P \\
	 \bb{\Delta} \bb{u}_N^* &=&
	 	\bb{R}_m \cdot \left(\bb{\Delta} \bb{u}\right)_P^*
\end{eqnarray}

Displacement increment values on the symmetry planes are calculated as
\begin{eqnarray}
	\left(\bb{\Delta} \bb{u}\right)_b = \frac{\left(\bb{\Delta} \bb{u}\right)_P^* + \left(\bb{\Delta} \bb{u}\right)_N^*}{2}
\end{eqnarray}
%where cell-centred values corrected for mesh skewness errors are

Further details on the implementation of symmetry planes in segregated cell-centred finite volume formulations can be found in \citet{Demirdzic2022}.

>>>>

%The two forms are connected through Nanson’s formula \cite{bathe_finite_1996}, which relate the deformed area vector $\bb{\Gamma}$ with the initial area vector $\bb{\Gamma}_{o}$:
%\begin{equation}
%    \bb{\Gamma} = J\bb{F}^{-T}\cdot\bb{\Gamma}_o
%\end{equation}
%Although the total Lagrangian approach is a viable option for wire drawing, the current work adopts the updated Lagrangian approach as developing Eulerian-type upstream and downstream conditions (Section \ref{sec:euler_BCs}) is conceptually easier in an updated Lagrangian formulation.

The governing equations are complemented by boundary conditions, with three types considered here: prescribed displacement, prescribed traction, and symmetry.
The definition of the engineering stress ($\bb{\sigma}_s$) and true stress ($\bb{\sigma}$) in Equations \ref{eqn:momentum_lingeom}, \ref{eqn:momentum_TL} and \ref{eqn:momentum_UL} is given by a chosen mechanical law.
Three mechanical laws are considered in this work, as briefly outlined in Appendix \ref{app:mechLaws}: linear elasticity (Hooke's law), neo-Hookean hyperelasticity, and neo-Hookean $J_2$ hyperelastoplasticity.

@@

%%--------------------------------------------------------------------------------------------------------------------%%
\subsection{Cell-Centred Finite Volume Discretisation}
\label{sec:discretisation}
%%--------------------------------------------------------------------------------------------------------------------%%
In this work, a nominally second-order cell-centred finite volume discretisation is employed.
% based on previously described work, for example, \citep{Cardiff2017, Batistic2022, Tukovic2013, Jasak2000}.
%Consequently, only a summary of the discretisation is presented below.
The solution domain is discretised in both space and time.
The total simulation period is divided into a finite number of time increments, denoted as $\Delta t$, and the discretised governing momentum equation is solved iteratively in a time-marching fashion.
%The spatial domain is partitioned into a finite number of contiguous convex polyhedral cells.
The spatial domain is partitioned into a finite set $\mathcal{P}$ of contiguous convex polyhedral cells, where each cell is denoted by $P \in \mathcal{P}$.
The number of cells in the mesh is indicated by $|\mathcal{P}|$.
A representative cell $P$ is shown in Figure~\ref{fig:cell}.
The set of all faces of cell $P$ is denoted by $\mathcal{F}_P$. This set is further subdivided into two disjoint subsets:
\begin{itemize}
    \item \textbf{Internal Faces} ($\mathcal{F}^{\text{int}}_P$): Faces that are shared with neighbouring cells.
    \item \textbf{Boundary Faces} ($\mathcal{F}^{\text{bnd}}_P$): Faces that lie on the boundary of the spatial domain.
    The boundary faces of cell $P$ are further classed into three disjoint sets, $\mathcal{F}^{\text{bnd}}_P \coloneqq \mathcal{F}^{\text{disp}}_P \cup \mathcal{F}^{\text{trac}}_P \cup \mathcal{F}^{\text{symm}}_P$, representing boundary faces where displacement ($\mathcal{F}^{\text{disp}}_P$), traction ($\mathcal{F}^{\text{trac}}_P$) and symmetry ($\mathcal{F}^{\text{symm}}_P$) conditions are prescribed.
       % \quad \text{with} \quad \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{trac}}_P = \mathcal{F}^{\text{disp}}_P \cap \mathcal{F}^{\text{symm}}_P = \mathcal{F}^{\text{trac}}_P \cap \mathcal{F}^{\text{symm}}_P = \emptyset.$
\end{itemize}
Each internal face $f_i \in \mathcal{F}^{\text{int}}_P$ corresponds to a neighbouring cell $N_{f_i} \in \mathcal{N}_P$, where $\mathcal{N}_P$ is the set of all neighbouring cells of $P$. The outward unit normal vector associated with an internal face $f_i$ is denoted by $\mathbf{n}_{f_i}$, while the outward unit normal vector associated with a boundary face $b_i$ is denoted by $\mathbf{n}_{b_i}$.
The vector $\mathbf{d}_{f_i}$ connects the centroid of cell $P$ with the centroid of the neighbouring cell $N_{f_i}$, whereas the vector $\mathbf{d}_{b_i}$ connects the centroid of cell $P$ with the centroid of boundary face ${b_i}$.
For convenience, we will also define the set of all faces of cell $P$ excluding those on a traction boundary as $\mathcal{F}_P^{\text{non-trac}} \coloneqq \mathcal{F}_P^{\text{int}} \cup \mathcal{F}_P^{\text{disp}} \cup \mathcal{F}_P^{\text{symm}}$.
%The proposed solution discretisation follows closely the approach of \citet{cardiff_lagrangian_2017}; consequently, only an overview of the final discretised form of equations and adopted solution algorithm are given below.
\begin{figure}[htbp]
	\centering
%	\subfigure[Magnitude of the manufactured displacement solution]
%	{
%		\label{fig:mms_solution}
   		\includegraphics[height=0.45\textwidth]{figures/cell} 
%   	}
%	\subfigure[Polyhedral mesh with $1\,000$ cells]
%	{
%		\label{fig:mms_mesh}
%   		\includegraphics[height=0.45\textwidth]{figures/mms_mesh}  
%   	}
	\caption{Representative convex polyhedral cell $P$ and neighbouring cell $N_{f_i}$, which share a face $f_i$.}
	\label{fig:cell}
\end{figure}

@@

Using Equation \ref{eq:volume_integral}, the inertia term (e.g. left-hand side term of Equation \ref{eqn:momentum_lingeom}) becomes
\begin{eqnarray} \label{eq:inertia}
	\int_{\Omega_P} \rho \frac{\partial \bb{u} }{\partial t}  d \Omega_P
%	\;&\approx&\;
%	\int_{\mathrm{\Omega}} \rho \left[ \frac{\partial \bb{u} }{\partial t}_P + (\bb{x} - \bb{x}_P) \cdot \nabla \frac{\partial \bb{u} }{\partial t}_P \right] d\mathrm{\Omega} \notag \\
	\;&\approx&\;
	\rho_P \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P  \Omega_P
\end{eqnarray}
Similarly, the body force term (e.g. the second term on the right-hand side of Equation \ref{eqn:momentum_lingeom}) becomes:
\begin{eqnarray}
	\int_{\Omega_P} \, \rho \, \bb{g} \,  d \Omega_P
	\;&\approx&\;
	\rho_P \, \bb{g}\,  \Omega_P
\end{eqnarray}
%where subscript $P$ indicates a quantity at the cell centre.
The discretisation of the acceleration in Equation \ref{eq:inertia} can be achieved using one of many finite difference schemes, e.g. first-order Euler, second-order backwards, second-order Newmark-beta.
In the current work, the second-order backwards (BDF2) scheme is used:
%\begin{eqnarray}
%	 \left(\frac{\partial^2 \bb{u} }{\partial t^2}\right)_P
%	 &\approx& \frac{3\bb{v}_{t+1} - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t} \notag \\
%	&\approx& \frac{3\left( \frac{3\bb{u}_{t+1} - 4\bb{u}_{t} + \bb{u}_{t-1}}{2\Delta t} \right) - 4\bb{v}_{t} + \bb{v}_{t-1}}{2\Delta t}
%\end{eqnarray}
\begin{eqnarray} \label{eq:inertia2}
	\left(\frac{\partial^2 \boldsymbol{u}_P}{\partial t^2}\right)_P
	&\approx& \frac{3\boldsymbol{v}_P^{[t+1]} - 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t} \notag \\
	&\approx&
	\frac{3\left( 
		\frac{3\boldsymbol{u}_P^{[t+1]} - 4\boldsymbol{u}_P^{[t]} + \boldsymbol{u}_P^{[t-1]}}{2\Delta t} 
		\right) 
	- 4\boldsymbol{v}_P^{[t]} + \boldsymbol{v}_P^{[t-1]}}{2\Delta t}
\end{eqnarray}
where $\Delta t$ is the time increment -- assumed constant here -- and $\bb{v} = \partial \bb{u}/\partial t$ is the velocity vector.
Superscript $[t]$ indicates the time level, with $\bb{u}_P^{[t+1]}$ corresponding to the unknown displacement at the current time step.
Consequently, the displacement and velocity at the two previous time steps must be stored, or alternatively the displacement at the previous four time steps.

@@

As noted, boundary conditions are enforced through the discretised surface integrals terms at the boundaries faces. % (Equations \ref{eq:divStressDiscret} and \ref{eq:RhieChow}).
If required, the displacement $\bb{u}_{t_i}$ on a traction boundary face $t_i$ can be calculated by extrapolation from the centre of cell $P$ as
\begin{eqnarray}
	\bb{u}_{t_i} = \bb{u}_P + \bb{d}_{t_i} \cdot \left(\bb{\nabla} \bb{u} \right)_P
\end{eqnarray}
where $\bb{d}_{t_i}$ represents the vector from the centroid of cell $P$ to the centroid of the traction boundary face $t_i$.

Similarly, if required, the displacement $\bb{u}_{s_i}$ at a symmetry plane face $s_i$ is calculated using the same approach as Equation \ref{eq:symm}:
\begin{eqnarray} 
	\bb{u}_{s_i}
		&=&  \frac{1}{2} \bb{u}_P + \bb{R}_{s_i} \cdot \bb{u}_P \notag \\
		&=& \left (\textbf{I} - \bb{n}_{s_i} \otimes \bb{n}_{s_i} \right) \cdot \bb{u}_P
\end{eqnarray}

@@

Within the square braces on thr right-hand side of Equation \ref{eq:RhieChow}, the first terms represent a compact stencil (two-node) approximation of the face normal gradient, while the second terms represent a larger stencil approximation.
These two terms cancel out in the limit of mesh refinement (or if the solution varies linearly); otherwise, they produce a stabilisation effect that tends to smooth the solution fields.
As the term reduces at a third-order \hl{or second?} rate, it does not affect the overall scheme's second-order accuracy; the 1-D local truncation error is derived for this Rhie-Chow stabilisation term in Appendix \ref{app:RhieChow}.

@@

%\subsubsection{Boundary Conditions}
%Boundary faces $f_i \in \mathcal{F}_{\text{bnd},P}$ are classified based on boundary conditions into displacement-type, traction-type, and symmetry-type faces.
%\hl{Comment on traction boundaries} \hl{extrapolate to get value or use constitutive law}
%Boundary conditions must be applied to the faces that coincide with the boundary of the solution domain.
%The discretised expressions on boundary faces are modified to account for either the known displacement components in Dirichlet conditions or the known traction for Neumann conditions.
%In the current approach, boundary face

%Three types of boundary condition are considered here: prescribed displacement, prescribed traction, and symmetry planes.
%
%Two terms must be approximated on a boundary face $b$:
%\begin{eqnarray} \label{eq:BCtermsLap}
%	\int_{\mathrm{\Gamma}_{b}} K_{imp} \bb{n}_{b} \cdot \left[ \nabla\left(\Delta\bb{u}\right) \right]_b d\mathrm{\Gamma}_{b}
%\end{eqnarray}
%which comes from Equation \ref{eq:diffusion}, and 
%\begin{eqnarray} \label{eq:BCtermsStress}
%	\int_{\mathrm{\Gamma_b}}(j_b \bb{f}^{-T}_b \cdot \bb{n}_b) \cdot \boldsymbol{\sigma}_b \ d\mathrm{\Gamma_b}
%\end{eqnarray}
%which comes from Equation \ref{eq:surfaceStress}, where subscript $b$ indicates a quantity at a boundary face in the updated configuration.
%
%Equation \ref{eq:BCtermsLap} must be discretised both implicitly and explicitly (deferred correction), as per Equation \ref{eqn:MomentumImplicitExplicit}, while Equation \ref{eq:BCtermsStress} is discretised explicitly.
%The Rhie-Chow term is set to zero on boundary faces, and all geometric terms ($\bb{n}_{b}$, $\mathrm{\Gamma}_{b}$) are calculated on the updated configuration.

%Five classes of boundary conditions are employed in the current work (prescribed displacement, prescribed traction, symmetry plane, axisymmetric/wedge, frictional contact) as described below.
%An additional boundary condition where the normal component of the displacement is prescribed and the shear traction is zero is also used; the discretisation of this condition employs the prescribed displacement condition in the boundary normal direction and the traction condition in the tangential direction.
%\begin{itemize}
%	\item Prescribed displacement,
%	\item Prescribed traction,
%	\item Symmetry,
%	\item Axisymmetric (wedge),
%	\item Frictional contact.
%\end{itemize}
%This enforcement of these five conditions is described below.

%\paragraph{Prescribed displacement}
%The displacement boundary condition, a Dirichlet condition, may be constant in time or time-varying and fixes the value of $\bb{u}$ at the centre of a boundary face.
%The corresponding displacement increment at the boundary face $\Delta \bb{u}_b = \bb{u}^{[m+1]}_b - \bb{u}^{m}_b$ is substituted into the calculation of the surface normal gradient term in Equation \ref{eq:BCtermsLap}.
%The resulting contribution of a boundary face $b$ to the discretised Laplacian term becomes:
%\begin{eqnarray} \label{eq:diffusionBC}
%%	\underbrace{
%	\int_{\mathrm{\Gamma}_{b}}
%	 K_{imp} \bb{n}_{b} \cdot \left[ \nabla\left(\Delta\bb{u}\right) \right]_b d\mathrm{\Gamma}_{b}
%%	 }_{\text {implicit}}
%	&\approx&
%	K_{imp}^b \left|\boldsymbol{\Delta}_{b} \right|\left(\frac{\Delta \bb{u}_{b} - \Delta \bb{u}_P}{\left|\bb{d}_b\right|}\right)\left|\bb{\Gamma}_{b}\right| \notag \\
%	    &&+ K_{imp}^b \; \bb{k}_{b} \cdot
%	    \left[
%	    \boldsymbol{\nabla} \left(\mathrm{\Delta}\bb{u}\right)
%	    \right]_P
%	    \left|\bb{\Gamma}_{b}\right|
%\end{eqnarray}
%where $\bb{d}_b$ connects the cell centre $\bb{C}_P$ with the boundary face centre $\bb{C}_b$ in the updated configuration, and the non-orthogonal correction vector $\bb{k}_{b} = (\textbf{I} - \bb{n}_b \bb{n}_b) \cdot \bb{d}_b$.

%\paragraph{Prescribed traction}
%The traction boundary condition, constant in time or time-varying, is implemented by replacing $\bb{n}_f \cdot \bb{\sigma}_f$ on boundary faces with the known prescribed traction $\bb{t}_b$.
%Following calculation of the cell-centred displacements, displacements on traction boundary faces $\bb{u}_b$ is calculated by extrapolation from the adjacent cell centre:
%\begin{eqnarray}
%	\bb{u}_b = \bb{u}_P + (\bb{x}_b - \bb{x}_P) \cdot \left(\bb{\nabla} \bb{u} \right)_P
%\end{eqnarray}
%where subscript $b$ indicates a value at the centre of a boundary face, and $P$ indicates a value at the centre of the adjacent cell.

%\paragraph{Symmetry plane}
%On a symmetry plane normal component of displacement is zero, and the tangential gradient of displacement is zero.
%For boundary faces on a symmetry plane, the mirror reflection $\left(\bb{\Delta} \bb{u}\right)_N$ of the cell-centre displacement increment $\left(\bb{\Delta} \bb{u}\right)_P$ can be obtained as
%\begin{eqnarray}
%	\left(\bb{\Delta} \bb{u}\right)_N &=& \bb{R}_m \cdot \left(\bb{\Delta} \bb{u}\right)_P \notag \\
%		&=& \left(\textbf{I} - 2 \bb{n}_b \bb{n}_b \right) \cdot \left(\bb{\Delta} \bb{u}\right)_P
%\end{eqnarray}
%where $\bb{R}_m$ is termed the reflection tensor.
%The discretisation in Equation \ref{eq:diffusion} takes the same form, but the neighbour cell-centre values are replaced by the mirror reflected values for boundary faces on the symmetry plane.
%The $\bb{d}_b$ vector at a symmetry plane boundary face is calculated as
%\begin{eqnarray}
%	 \bb{d}_b &=& \bb{R}_m \cdot \bb{C}_P - \bb{C}_P
%\end{eqnarray}
%%where $\bb{C}_P$ is the positional vector of the centre of cell $P$.
%This means that symmetry planes are orthogonal by definition, and the non-orthogonal correction (second term on the right-hand side of Equation \ref{eq:diffusion}) drops out.
%
%In the current work, the cell-centre values are corrected for skewness errors, resulting in the following contribution to the Laplacian term at a symmetry boundary face (Equation \ref{eq:diffusionBC}):
%\begin{eqnarray}
%	\int_{\mathrm{\Gamma}_{u_b}}	 K_{imp} \bb{n}_{b} \cdot\nabla\left(\Delta\bb{u}\right) d\mathrm{\Gamma}_{b}
%	&=&
%	K_{imp}^b \left|\boldsymbol{\Delta}_{b} \right| \left( \frac{\bb{\Delta} \bb{u}_N^* - \Delta \bb{u}_P^*}{\left|\bb{d}_b\right|} \right) \left|\bb{\Gamma}_{b}\right|
%\end{eqnarray}
%where the skew-corrected cell-centre values are
%\begin{eqnarray}
%	 \bb{\Delta} \bb{u}_P^* &=&
%	 	\bb{\Delta} \bb{u}_P + \left[ (\textbf{I} - \bb{n}_b \bb{n}_b) \cdot \bb{d}_b \right] \cdot \left[ \bb{\nabla} \left(\bb{\Delta} \bb{u}\right) \right]_P \\
%	 \bb{\Delta} \bb{u}_N^* &=&
%	 	\bb{R}_m \cdot \left(\bb{\Delta} \bb{u}\right)_P^*
%\end{eqnarray}
%
%Displacement increment values on the symmetry planes are calculated as
%\begin{eqnarray}
%	\left(\bb{\Delta} \bb{u}\right)_b = \frac{\left(\bb{\Delta} \bb{u}\right)_P^* + \left(\bb{\Delta} \bb{u}\right)_N^*}{2}
%\end{eqnarray}
%%where cell-centred values corrected for mesh skewness errors are
%
%Further details on the implementation of symmetry planes in segregated cell-centred finite volume formulations can be found in \citet{Demirdzic2022}.
